---
title: 'autocovarianceTesting: Test for Equality of Autocovariance Functions in Time
  Series'
author: "Daniel Cirkovic, Thomas J. Fisher "
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The package 'autocovarianceTesting' implements multiple hypothesis tests for testing equality of autocovariance functions for two mean zero, stationary time series. The time series are allowed to be multivariate and linearly dependent. We implement methods assuming independence of the series from ["Testing equality of stationary autocovariances." Lund et. al (2009)](https://doi.org/10.1111/j.1467-9892.2009.00616.x) and new extensions to linearly dependent times series. We also implement a weighting mechanism to the former tests by borrowing ideas from ["New Weighted Portmanteau Statistics for Time Series Goodness of Fit Testing" Fisher and Gallagher (2012)](https://doi.org/10.1080/01621459.2012.688465). The package also implements bootstrap methods from ["A computational bootstrap procedure to compare two dependent time series" Jin et. al (2019)](https://doi.org/10.1080/00949655.2019.1639704), offering extensions to multivariate series. 

'autocovarianceTesting' includes one function: `autocovarianceTest`. `autocovarianceTest` allows the the user to supply two time series and select from any combination of the previously mentioned methods to test for equality of autocovariance functions. Choices of methods are made through the `test` argument. See function documentation for more details.

## Installation

The package is currently available on github. To install, use the following:

```{r, eval = FALSE}
# Installation of autocovarianceTesting
devtools::install_github("cirkovd/autocovarianceTesting", build_vignettes = TRUE)
```

## Example

For quarterly change in US consumption and income from 1970 to 2016, test if the autocovariance structure of the series differ up to lag 12.

```{r, warning = FALSE, message = FALSE}
library(autocovarianceTesting)
library(fpp2) # not needed for package, just used for data
head(uschange)

# Structure data
t <- time(uschange)
# US consumption
usconsumption <- matrix(uschange[ ,1])
# US income
usincome <- matrix(uschange[ ,2])

# Plot the two datasets, no obvious skewness, minor outliers
plot(t, usconsumption, type = "l", xlab = "Time", ylab = "Consumption")
plot(t, usincome, type = "l", xlab = "Time", ylab = "Income")

# The series exhibit clear dependence
ccf(c(usconsumption), c(usincome), main = "Cross-Correlation Function")
```

Since the series exhibit clear correlation, only use the dependent tests provided in `autocovarianceTest`.

```{r}
# Make the series mean zero
usconsumption <- usconsumption - mean(usconsumption)
usincome <- usincome - mean(usincome)

set.seed(4321)
# Run test for up to lag 12
output <- autocovariance_test(usconsumption, usincome, max_lag= 12, 
                   test = c("Fixed_Lag", "Weighted_Fixed_Lag", 
                            "Auto_Lag_Jin", "Auto_Lag_Bartlett"),
                   num_bootstrap = 1000, prewhiten = TRUE)
output
```

All tests reject the null hypothesis of equality of autocovariances up to lag 12.

```{r}
plot(output)
```

For more on assumption checking and general function use, please see the vignette. 
